---
title: "A&E Attendances Dataset"
date: "`r Sys.Date()`"
date-format: "D MMMM YYYY"
format:
  html:
    # Table of Contents options
    toc: true
    toc-depth: 3
    toc-location: left
    toc-title: Contents
    number-sections: false
    number-depth: 3
    # Render options
    theme: cosmo
    anchor-sections: false
    html-math-method: katex
    # Code options
    code-tools:
      source: false
      toggle: false
      caption: none
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    standalone: true
    # URL options
    link-external-icon: true
    link-external-newwindow: true
    # References
    citations-hover: true
    footnotes-hover: true
    callout-appearance: simple
    callout-collapse: true
    cap-location: bottom
    title-block-banner: '#005EB8'
    backgroundcolor: '#f0f4f5'
    mainfont: 'Open Sans'
crossref:
  fig-title: '**Figure**'
  fig-labels: arabic
#execute: 
#  echo: false
---

```{r load library}
#| include: False

# Load libs and global vars
# -------------------------------------------------------------------------
suppressPackageStartupMessages({
  library(targets)
  library(dplyr)
  library(readr)
  library(here)
})
#library(here)
#here()
#library(ggrepel)
#library(tidyverse)
#options(tidyverse.quiet = TRUE)
#library(dplyr)
#library(lemon)
#library(sf)
##library(scales)
#library(zoo)
#library(lubridate, warn.conflicts = FALSE)
#library(plyr, include.only = c("rbind.fill"))
#library(plotly)
#library(DT)

# Needed
#library(scales)
#library(ggrepel)
#library(lubridate)
#library(dplyr)
#library(forcats)
#library(tidyr)
```

```{css, echo=FALSE}
.panel-tabset .nav-item {
  font-size: 16px;
}

.dt-buttons {
    /* margin-bottom: -7px; */
    position: absolute;
    bottom: 0;
    left: 0;
    right: 0;
    text-align: center;
    padding-bottom: 0.3em;
}

.figure-caption {
  color: #373a3c;
  font-size: 16px;
}

.nav-tabs .nav-link.active, .nav-tabs .nav-item.show .nav-link {
  color: #495057;
  background-color: #fff;
  border-color: #dee2e6 #dee2e6 #fff;
}

.tab-content {
  padding-top: 1em;
  padding-bottom: 1em;
  background-color: #fff;
}

.quarto-title-banner {
  padding-top: 30px;
  padding-right: 0px;
  padding-bottom: 40px;
  padding-left: 0px;
  margin-bottom: 0px;
}

.quarto-title-meta  {
  padding-top: 0px;
  padding-right: 0px;
  padding-bottom: 0px;
  padding-left: 0px;
  margin-bottom: 0px;
  background-color: #fff;
}

#title-block-header.quarto-title-block {
  padding-top: 0px;
  padding-right: 0px;
  padding-bottom: 10px;
  padding-left: 0px;
  margin-bottom: 0px;
  background-color: #fff;
}

.quarto-title {
	background-image: url("data:image/svg+xml,%3Csvg xmlns='http://www.w3.org/2000/svg' viewBox='0 0 40 16'%3E%3Cpath fill='%23fff' d='M0 0h40v16H0z'%3E%3C/path%3E%3Cpath fill='%23005eb8' d='M3.9 1.5h4.4l2.6 9h.1l1.8-9h3.3l-2.8 13H9l-2.7-9h-.1l-1.8 9H1.1M17.3 1.5h3.6l-1 4.9h4L25 1.5h3.5l-2.7 13h-3.5l1.1-5.6h-4.1l-1.2 5.6h-3.4M37.7 4.4c-.7-.3-1.6-.6-2.9-.6-1.4 0-2.5.2-2.5 1.3 0 1.8 5.1 1.2 5.1 5.1 0 3.6-3.3 4.5-6.4 4.5-1.3 0-2.9-.3-4-.7l.8-2.7c.7.4 2.1.7 3.2.7s2.8-.2 2.8-1.5c0-2.1-5.1-1.3-5.1-5 0-3.4 2.9-4.4 5.8-4.4 1.6 0 3.1.2 4 .6'%3E%3C/path%3E%3C/svg%3E");
      display: block;
      background-color: #005EB8;
      background-repeat: no-repeat;
      background-size: 90px;
      background-position: 0px 0px;
}

.quarto-title p.subtitle {
      margin-top: 10px;
      font-size: 22px;
      font-weight: 300;
}

.quarto-title h1.title {
      text-indent: 3.7em;
      font-size: 30px;
}

.callout-caption-container {
  color: #005EB8;
}

.callout {
  background-color: white;
  margin-top: 0.5rem;
  margin-bottom: 0.5rem;
  border-radius: .25rem;
}

.callout-note.callout {
  border-left-color: #005EB8;
  background-color: white;
}

.callout.callout-style-simple {
  padding: .01em .7em;
}

.title {
  font-size: 28px;
  font-weight: lighter;
}

h1, .h1 {
  font-size:30px;
  font-weight:600;
}

h2, .h2 {
 font-size:20px;
 font-weight:600;
}

h3, .h3 {
 font-size:18px;
 font-weight:600;
}

.header-section-number {
  color: #373a3c;
}

@import url('https://fonts.googleapis.com/css?family=Open+Sans');

.opensans {
  font-family: "Open Sans", sans-serif;
}

body {
 -moz-osx-font-smoothing:grayscale;
 -webkit-font-smoothing:antialiased;
}
```

# The A&E Attendances Dataset

This report explains how to use the `ae_attendances` dataset from the [{NHSRdatasets}](https://nhs-r-community.github.io/NHSRdatasets/) package. The dataset contains all reported A&E attendances for the period April 2016 through March 2019.

First let's open and explore the dataset to show the first 10 rows of data.

:::: panel-tabset

## A&E Attendances

```{r data table}
# load dataset and create formatted table
# -------------------------------------------------------------------------
tar_load(ae_attendance_raw)

ae_attendance_raw %>% 
  head(10) %>%
  knitr::kable()
```

## Reference

::: {.callout-note collapse="false"}
## Data source

- [NHS England Statistical Work Areas][nhs_swa]
- available under the [Open Government Licence v3.0][ogl_v3]
:::

The dataset contains:

- `period`: the month that this activity relates to, stored as a date (1st of each month)
- `org_code`: the [ODS][ods]
code for the organisation that this activity relates to
- `type`: the [Department Type][ae_types]
for this activity, either 1, 2, or other
- `attendances`: the number of attendances for this department type at this organisation
for this month
- `breaches`: the number of attendances that breached the 4 hour target
- `admissions`: the number of attendances that resulted in an admission to the
hospital

:::: 

# Footnotes

Adapted from the vignette in the [{NHSRdatasets}](https://nhs-r-community.github.io/NHSRdatasets/) package, writen by [Thomas Jemmett][tj_email] and developed by the [NHS-R Community](https://nhsrcommunity.com/).

[tj_email]: mailto:thomas.jemmett@nhs.net
[su_web]:   https://www.strategyunitwm.nhs.uk/ 
[nhs_swa]:  https://www.england.nhs.uk/statistics/statistical-work-areas/ae-waiting-times-and-activity/
[ogl_v3]:   https://www.nationalarchives.gov.uk/doc/open-government-licence/version/3/
[ods]:      https://digital.nhs.uk/services/organisation-data-service
[ae_types]: https://web.archive.org/web/20200128111444/https://www.datadictionary.nhs.uk/data_dictionary/attributes/a/acc/accident_and_emergency_department_type_de.asp